Modular Synthesis of Verified Verifiers of Computation with STV Algorithms

### 2019-01-26 paper/paper.tex
introduction

\documentclass[10pt,conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amssymb,amsfonts}
%amsmath,
\usepackage{mathtools}

@@

\maketitle
\begin{abstract}
\end{abstract}

@@

\end{IEEEkeywords}
\section{Introduction}\label{sec:intro}
%\setlength{\abovedisplayskip}{0.5em}
%\setlength{\belowdisplayskip}{0.5em}
\section{Evidences}
\label{sec:DataEv}
%\section{The Generic STV Algorithm}
%\label{sec:GenSTV}
%%%%We shall initially clarify what the problem is that we are trying to address. 
%Then we proceed to elaborate on our approach and detail about the solution. 
%%%%%  To this end assume an arbitrary scheme $\mathcal{A}$ belongs to the STV family.The question is given any imp $\mathcal{P}$ is any implementation of $\mathcal{A}$ where the source code of $\mathcal{P}$ is  secret. 
%The question is how can one verify  $\mathcal{P}$ implements $\mathcal{A}$ correctly so that computation carried out by $\mathcal{P}$ is trustworthy?
%%%%The question is how can one verify that each execution of $\mathcal{P}$ correctly computes the end result in accordance with instructions in $\mathcal{A}$?

>>>>

\documentclass[10pt,conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{wrapfig}
\usepackage{amssymb,amsfonts}
%amsmath,
\usepackage{mathtools}

@@

\maketitle
\begin{abstract}
Single transferrable vote (STV) is a family of preferential voting systems, different instances of which are used in binding elections throughout the world. We give a formal specification of this family, from which we derive fully verified tools that verify the computation for various instances of STV vote counting. These tools validate the probably correct execution of a run of a vote counting algorithm, based on a transcript of the count.

Our framework distils the similarities and differences of various instances of STV and gives a uniform and modular way of synthesising verifiers for its various instances, and provides the flexibility and ease for adapting and extending it to a variety of STV schemes. We minimise the trusted base in correctness of the tools produced by using the HOL4 and CakeML as the technical basis. We first formally specify and verify the tools in HOL4 and then obtain the machine executable versions for the tools by relying on the verified proof translator and the compiler of the CakeML. Moreover, proofs that we establish in HOL4 and CakeML are almost completely automated so that new verified instances of STV can be created with no (or minimal) extra proof. Finally, our experimental results with executable code demonstrate feasibility of deploying the framework for verifying real size elections having an STV counting algorithm.
\end{abstract}

@@

\end{IEEEkeywords}
\section{Introduction}\label{sec:intro}

Single transferable vote (STV) \footnote{DP: I vaguelu reall that
there's a book that speaks about STV, at least in part? A reference
would be good here} is a preferential voting scheme used
in various jurisdictions around the world, including Ireland, Malta,
India, Pakistan, Australia, New Zealand and some municipialities of
the United States, typically for multi-seat constituencies. 
A ballot is a preference-ordered list of
candidates, typically obtained specifying a numerical preference
for each candidate appearing on the ballot paper as illustrated in Figure
\footnote{DP: remove figure and explain textually if low on space}
\ref{fig:ballot}.

\begin{figure}\label{fig:ballot}
\begin{center}
\includegraphics[width=0.18\textwidth]{bal-cropped.pdf}
\end{center}
\caption{A hypothetical STV ballot}
\end{figure}

Each specific STV scheme first determines a \emph{quota}, i.e. the
(minimal) number of votes that a candidate needs to attract to be
elected (often the Droop quota \footnote{DP: reference}). The quota
depends on the number of ballots cast, and the number of vacancies
to fill. Once the quota has been determined, tallying proceeds
as follows:
\begin{enumerate}
  \item count all first preference votes
  \item if a candidate's first preference votes are above the quota,
  this candidate is elected
  \item if an elected candidate has received votes in excess of
  the quota, these  excess votes are transferred to the next preference
  (possibly resulting in more candidates being elected)
  \item if there are still vacancies to fill and no candidate is
  being elected, the candidate with the least number of first
  preference votes is eliminated, and their votes are transferred to
  the next preference (possibly resulting in more candidates being
  elected)
  \item this process continues until either all vacancies are
  filled, or the number of elected candidates, together with the
  candidates not yet eliminated matches the number of vacancies.
\end{enumerate}

This informal description is necessarily imprecise, with details
differing widely between jurisdictions. For example,
the details of precisely which excess ballots are being transferred
after candidates have been elected, the order in which surplus
ballots are being transferred vary widely between different
jurisdictions.

The goal of this paper is to advance formal methods in elections by
demonstrating how STV vote counting can be verified in a modular,
transparent, and practical feasible way over a minimal computational
trust base:  we show how to synthesise practically feasible and provably correct verifies for a
large class of STV schemes in a modular and almost fully automated
way.

%\texttt{Possibly explain transparent and trust base?}
%Goals.
%\begin{itemize}
%  \item verification of computation of an entire family of STV
%  algorithms
%  \item advance the use of formal methods in elections
%  \item minimize the trust base 
%  \item practical feasibility (development / runtime)
%  \item transparency
%\end{itemize}

We achieve this by first giving a formal definition STV (called
\emph{generic STV}) 
that 
encompasses the entire family of STV vote counting
schemes. This definition necessarily under-specifies any particular
STV vote counting scheme, but fixes both basic data structures and
mode of computation common to all instances of STV. Trustworthyness is achieved by
not verifying the \emph{software} that computes the result, but
instead by verifying a \emph{certificate} that substantiates the
correctness of the count. A certificate records the individual
steps taken in the computation of winners. (That is, our approach is
an instance of verifiable computation \footnote{DP: reference}). The
certificate directly corresponds to the individual steps of the
count (and could so also be verified, or spot-checked manually),
giving transparency. We minimize the computaional trust base by
verifying our executables down to the machine level using a
combination of HOL \footnote{DP: reference} and CakeML \footnote{DP:
reference}.

We observe that every instance of STV performs the same \emph{actions}
(e.g. count first preferences, transfer a surplus, eliminate a
candiate, etc.) where the precise interpretation of action differs.
We formalise these actions as rules that advance the state of the
count towards a final result. A certificate then becomes a
sequence of rule applications. While generic STV abstracts the
communalities, each \emph{instance} of STV formally specifies the
pre- and postconditions for each rule as well as a boolean function
that decides whether a rule has been applied correctly.

Executable certificate verifiers are the constructed almost
automatically: 
almost all of the ensuing proof obligations are discharged
generically, and we fully 
automate the subsequent translation into machine-level verified executable using CakeML.

%Approach.
%\begin{itemize}
%  \item formal definition of the family of STV algorithms
%  \item trustworthyness through external (certificate) validation
%  \item verification of certificates, rather than of implementations
%  \item certificates are rule applications
%  \item trustworthyness through trustworthy certificate verifiers
%  \item minimal computational trust base through  cakeml and hol
%  \item practical feasibility through modularity and proof
%  automation
%  \item full automation of of cakeml translation
%  \item automation of discharge of proof obligations
%\end{itemize}

We evaluate our framework by analysing the degree of automation and
modularity achieved, and by giving realistic benchmarks results over
real-world eletion data.

In summary, we present a framework that allows us to formally
specify instances of STV and sythesise an executable
certificate verifier in a modular way that is itself verified down
to the machine level.
%Evaluation.
%\begin{itemize}
%  \item minimal trust base realised throug cakeml and hol
%  \item aplicability through benchmarks
%  \item adaptability: combinnation of rules are automatic
%  \item transparency through certificate format
%\end{itemize}

%\setlength{\abovedisplayskip}{0.5em}
%\setlength{\belowdisplayskip}{0.5em}
\section{Evidences}
\label{sec:DataEv}
%\section{The Generic STV Algorithm}
%\label{sec:GenSTV}
%%%%We shall initially clarify what the problem is that we are trying to address. 
%Then we proceed to elaborate on our approach and detail about the solution. 
%%%%%  To this end assume an arbitrary scheme $\mathcal{A}$ belongs to the STV family.The question is given any imp $\mathcal{P}$ is any implementation of $\mathcal{A}$ where the source code of $\mathcal{P}$ is  secret. 
%The question is how can one verify  $\mathcal{P}$ implements $\mathcal{A}$ correctly so that computation carried out by $\mathcal{P}$ is trustworthy?
%%%%The question is how can one verify that each execution of $\mathcal{P}$ correctly computes the end result in accordance with instructions in $\mathcal{A}$?
